\chapter{Conclusions}
\label{chap:conclusions-en}

This thesis began with an open question, whether human discussion facilitation can be described clearly enough to be implemented as a system. The literature made it possible to separate the process into observable decisions, and the implementation showed that these decisions can be organized as a pipeline. The resulting artifact receives a thread, analyzes its state, decides whether to intervene, and generates a facilitation proposal when appropriate.

\section{Achievement of the objectives}

The main objective was to design and implement an engine based on LLMs and context-engineering techniques for analyzing and facilitating asynchronous academic discussions from different sources. This objective was achieved within its design and implementation scope, the artifact uses thread content, course context, and literature-derived criteria to decide whether to intervene and to produce a contextualized response. The input contract also separates the format processed by the engine from the educational platform that supplies the discussion. A source such as Open edX can therefore be replaced by another LMS or by local files through an adapter without changing the pipeline.

The experiments provide evidence of technical feasibility for that objective. The architecture operated end to end with local and external models. Facilitation was designed for asynchronous discussions, in which participants post at different times and the system's intervention does not need to appear during a real-time conversation. This characteristic makes the processing times observed in the experiments tolerable. Adopting an asynchronous flow was also a technical decision because it reduced latency and concurrency requirements and allowed the work to focus on analyzing the thread and constructing the intervention.

This evidence does not yet answer the underlying educational question. The system can generate an intervention intended to support a student, but the study did not establish whether it helps an instructor maintain teaching quality as the number of students or threads grows. The engine was designed, implemented, and shown to operate; pedagogical improvement and scalability of teaching were not established.

The first specific objective concerned a staged agent system. Dividing the problem worked well during design and implementation. Each agent contains one responsibility, its own context, instructions, and output schema. This separation made a non-deterministic task easier to reason about. It also allowed, for example, the classifier's instructions and output schema to change without modifying the agents that decide whether to intervene or draft the response. The pipeline, discussion backend, and storage layer were also designed as replaceable components.

No single-prompt baseline was run against the same models. The evidence therefore does not show that the multi-agent architecture produces better interventions than a direct model call. Its observed benefit in this project was an engineering one: responsibilities became explicit, intermediate decisions could be inspected, and components could be modified independently.

The second objective was to define a literature-grounded context-engineering strategy. Facilitation roles became the starting point for the design. Studying how a human facilitator intervenes made it possible to translate pedagogical knowledge into responsibilities, criteria, and techniques. This translation made the system implementable, but it remains an engineering interpretation. Educational specialists and classroom experience are required before it can be considered a validated pedagogical strategy.

The source-independence objective was also met as a technical demonstration. The common contract supported reproducible scenarios, six anonymized historical threads preserved in the repository, and a partial Open edX source without coupling the pipeline to a single data representation. The source, storage layer, and pipeline organization can be replaced through their interfaces. The Open edX work demonstrates data exchange, not routine use within a live course.

The final objective concerned system evaluation. The two campaigns recorded executions with local and external models and produced 46 scoreable interventions. The LLM judge assigned a mean score of 3.911 out of 5 and marked 29 cases for review, including 27 with information unsupported by the thread. These data provide an initial view of output quality and failure modes, but the judge is itself a general-purpose model. Its scores do not replace expert judgment or establish pedagogical effectiveness.

\section{Lessons from design and implementation}

The six milestones in the introduction were completed with different levels of scope. Problem definition and literature review described the process; implementation produced the engine and dashboard; the common contract and Open edX addressed discussion sources; the Idril and EC2 campaigns provided preliminary technical evaluation; and both environments were deployed for experimentation and demonstration.

The most difficult part was understanding model behavior and building a system around it. Input form, structured-output mode, model family, and model size affected the results. Some failures were transient, while others appeared only with particular local models. The system therefore needed tracing, time limits, partial-result handling, and model-specific output extraction in addition to its logical workflow.

The two environments represented different operating conditions. External models were faster in these runs, but sustained use may be difficult within a university budget. Idril represented an institution-hosted alternative and exposed compatibility and structured-output cases that did not appear in the same way with external models. The thesis does not prescribe either option; it documents that the same architecture can use both.

\section{Contribution}

The main contribution is the architecture, a modular and traceable implementation for studying LLM-based facilitation of asynchronous academic discussions. It turns facilitation roles into separate responsibilities, preserves intermediate decisions, and allows the data source, storage layer, models, and pipeline organization to be replaced.

The artifact does more than generate text. It provides a way to inspect how the decision to intervene was reached. This is particularly relevant in a non-deterministic process, where schema-valid output may still contain an inappropriate decision or unsupported claim.

\section{Limitations}

\begin{itemize}
	\item The system was not evaluated with students, instructors, or operating courses. There is therefore no evidence yet that its interventions are useful in real settings, reduce instructor workload, or allow more discussions to be supported without loss of quality.
	\item Quality assurance covered two campaigns, eight model configurations, and 18 threads. This coverage exposed failures but does not constitute exhaustive testing across other disciplines, thread lengths, group sizes, languages, or model behaviors.
	\item The state taxonomy and facilitation roles were derived from the literature and specified from an engineering perspective. They were not validated by specialists, and a single label may simplify discussions that combine several states or evolve in an unanticipated way.
	\item LLM-as-judge assessment has not been calibrated against formal expert annotation.
	\item There is no controlled comparison between the multi-agent pipeline and a single-prompt baseline.
	\item Local models presented compatibility and structured-output problems, and model-reported confidence was not calibrated against observed outcomes.
	\item The pipeline includes the \texttt{instructor\_escalation} technique, prevents instructor notes from being posted to the thread, and can flag content for review. Tests exercised the technical path and the experiments produced escalation cases, but instructors did not validate whether the correct cases were referred or whether the generated information was sufficient to resolve them.
\end{itemize}

\section{Future work}

The next stage should incorporate expert pedagogical knowledge. Educational specialists should review the roles, intervention policies, and prompts using both the literature and practical classroom experience. Their participation would support an annotation protocol, calibration of automatic judging, and review of cases containing unsupported information.

The need to intervene should be validated separately from the quality of the generated text. Several instructors could independently review a sample of threads and decide whether to intervene, wait, or escalate. Comparing those decisions with the pipeline would reveal unnecessary interventions, missed cases, and situations that do not fit the current taxonomy.

Future work on escalation starts from the mechanism already implemented. Operational validation with instructors should examine which situations ought to be referred, whether the summary contains the information needed, how the notification fits their workflow, how long resolution takes, and how the outcome is recorded. This test would help adjust thresholds and distinguish a useful alert from one that merely adds to instructor workload.

After calibration, the artifact should be tested in real courses. Evaluation should consider both students and instructors: intervention quality, discussion development, learning outcomes, instructor workload, and the number of threads that can be supported without reducing teaching quality. Only then can the system's contribution to educational scalability be studied.

Further work should also complete a live LMS integration, add a single-prompt baseline, and compare local and external configurations using latency, success-rate, and cost metrics. A simulation environment with student agents representing different knowledge levels and participation styles could provide an intermediate stage before trials with students \parencite{ZhangEtAl2024simclass}.

In summary, this thesis demonstrates the technical feasibility of a modular architecture for asynchronous academic discussion facilitation. The architecture receives, processes, and responds; it can use local or external models and does not depend on a single data source. The decisive question remains open, whether this support improves the student experience and allows instructors to manage more discussions without reducing teaching quality.
